---
title: "Week-4"
author: "kevinoliva26"
date: "7/7/2020"
output: html_document
---

###str function

Da los argumentos si se pone en una funcion, o da informacion de algun data type

```{r}
str(airquality)
```

###generating random numbers
Each probability distribution in R has an r*** function (for "random"), a d*** function (for "density"), a p*** (for "probability"), and q*** (for "quantile"). 

rnorm(x): generates random variables with a given mean an sd
dnorm(q): evaluate the normal probability densitry at a point
pnorm(p): evaluate the cumulative distribution for a normal distribution
qnorm(p): inversa de pnorm 
rpois(n): generates random Poisson variables with a given rate
ppois: distrinusion acumulada de Poisson



These functions have usually 4 functions associated with them:

d density
r random number generation
p cumulative distribution
q quantile function

```{r}
#Para volver a generar los mismos numeros aleatorios. set.seed(1) indica que voy a volver a generar los mismos numeros aleatorios que la primera vez que lo genere. 
set.seed(1)
rnorm(5)
```

###simulating a linear model

```{r}
set.seed(20)
x <- rnorm(100)
e <- rnorm(100,0,2)
y <- 0.5 + 2 * x + e
summary(y)
plot(x,y)
```

Que pasa si x es binario

```{r}
set.seed(10)
x <- rbinom(100, 1, 0.5)
e <- rnorm(100,0,2)
y <- 0.5 + 2 * x + e
summary(y)
plot(x,y)
```

###random sampling

The sample funcionts draw randomly from a specified vector of objects allowing you to sample from arbitrary distributions

```{r}
set.seed(1)
sample(1:10,4)
sample(letters,5)
sample(1:10, replace = TRUE)
```

##R Profiler

Es una herramienta de optimizacion de codigo, como por ejemplo cuando una porcion del codigo tiene tiene que repetirse miles de veces y eso hace que tarde en correr. 

system.time() sirve para saber cuanto tarda en correr una expresion. regresa un objeto de classe proc_time

```{r}
system.time(readLines("http://www.jhsph.edu"))
```


Rprof() funcion para iniciar el R Profiler
summaryRprof() resume el output de Rprof para que sea mas readable

Keeps track of the function call stack at reagularly sampled intervals and tabulates how much time is spend in each function.








