---
title: "Week-3"
author: "kevinoliva26"
date: "7/2/2020"
output: html_document
---

## Loop Functions

Tienen la palabra apply en su nombre

lapply: loop over a list and evaluate a function on each element
sapply: same as lapply but try to simplify the result
apply: apply a funcion over the margins of an array
tapply: apply a funciont over subsets of a vector
mapply: multivariate version of lapply

###lapply()
Always return a list
```{r}
x <- list(a=1:4,b=rnorm(10),c=rnorm(20,1),d=rnorm(100,5))
lapply(x,mean)
```

```{r}
x<- 1:4
lapply(x,runif,min=0,max=10)
```

###sapply()
Always return a list length 1

```{r}
x <- list(a=1:4,b=rnorm(10),c=rnorm(20,1),d=rnorm(100,5))
sapply(x,mean)
```
if the result is a list where every element is of length one, then sapply() returns a vector. If the result is a list where every element is a vector of the same length (> 1), sapply() returns a matrix. If sapply() can't figure things out, then it just returns a list, no different from what lapply() would give you.



###apply()
Aplica una funcion ya sea a las columnas filas o columnas seleccionadas. 
is used with arrays
is often used to apply a function to the rows or columns of a matrix
argument MAGIN is an integer vector indicating which margins should be retained
... quiere decir que se pasan los argumentos al final de apply() hacia la funcion FUN
```{r}
x <- matrix(rnorm(200),20,10)
apply(x,2,mean)
apply(x,1,sum)
```
Con el 1, se conservan las filas
Con el 2, se conservan las columnas

rowSums = apply(x, 1, sum)
rowMeans = apply(x, 1, mean)
colSums = apply(x, 2, sum)
colMeans = apply(x, 2, mean)

```{r}
x <- matrix(rnorm(200),20, 10)
apply(x, 1, quantile, probs = c(0.25,0.75))
```

###vapply
Similar to sapply but requires you to specify the format of the output in advance, instead of just allowing R to 'guess' what you wanted.
```{r}
vapply(flags, class, character(1))
```


###mapply
Applies a function over multiple lists arguments
function (FUN,...,MoreArgs=NULL,SIMPLIFY=TRUE)
FUN: Function to apply
... arguments to apply over
MoreArgs: a list of other arguments to FUN
SIMPLIFY: indicates wheter the result should be simplified

```{r}
mapply(rep,1:4,4:1)
```


###tapply()
is used to apply a funciont over subsets of a vector.
Acepta subsets como argumentos
Se parece a la funcion de group by de SQL
function(x, INDEX, FUN = NULL,..., simplify = TRUE)
x: vector
INDEX: a factor or a list of vector
FUN: function to be applied
... contains other arguments to be passed to FUN
simplify, should we simplefy the result?

```{r}
library(datasets)
data(iris)
tapply(iris$Sepal.Length, iris$Species, mean)
```


```{r}
x<- c(rnorm(10),runif(10),rnorm(10,1))
f <- gl(3,10)
f

tapply(x, f, mean)
```


```{r}
x<- c(rnorm(10),runif(10),rnorm(10,1))
f <- gl(3,10)
f

tapply(x, f, mean,simplify = FALSE)
```


##split()
Takes a vector and splits it into groups 
function(x, f, drop= FALSE,...)
x: vector of data frame
f: factor (or coerced to one) or a list of factor
drop: indicates wheter empty factor levels should be dropped
```{r}
x <- c(rnorm(5),runif(5))
x
f <- gl(2,5)
split(x,f)
```


```{r}
library(datasets)
head(airquality)
```

Splitting data into monthly pieces with lapply()
```{r}
library(datasets)
s <- split(airquality, airquality$Month)
lapply(s, function(x) colMeans(x[,c("Ozone","Solar.R","Wind")]))
#Function(x) es una mini funcion que defino inmediatamente. 
```

Splitting data into monthly pieces with lapply()
```{r}
sapply(s, function(x) colMeans(x[,c("Ozone","Solar.R","Wind")]))

sapply(s, function(x) colMeans(x[,c("Ozone","Solar.R","Wind")],na.rm=TRUE))
```


##Debugging Tools

message: notificacion producida por una ejecucion
warning: notificacion the que algo esta mal pero no es fatal. la ejecucion sigue.
error: error fatal que para la ejecucion.
condition: algo inesperado que sucede. programadores pueden crear sus condiciones

invisible(x) <- no imprime x en una funcion

###Primary tools for debuging

debug: gives a funcion a flag, and allows you to step trhough execution
traceback: prints out the function call stack after an error occurs. Does nothing if there is no error
recover: allows you to modify the error behavior so that you can browse the function call stack
browser: suspends the execution of a function wherever it is called and puts the function in debug mode
trace: allows you to insert debugging code into a function a specific places

